The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2016-11-30 manuscript/manuscript.Rnw
clean up the abstract

\begin{abstract}
  \textbf{Introduction} The placenta is arguably the most
  anatomically variable organ in mammals.

@@

  \textbf{Results} We identified a set of
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  core genes which are expressed (FPKM $\ge 10$) in all placentas, %
  %
  including genes with immune-modulating properties (\ac{ANXA2},
  \ac{ANXA1}, \ac{S100A11}, \ac{S100A10}, and \ac{LGALS1}), cell-cell
  interactions (\ac{LAMC1}, \ac{LUM}, and \ac{LGALS1}),
  invasion (\ac{GRB2} and \ac{RALB}) and syncytialization
  (\ac{ANXA5} and \ac{ANXA1}). %
  We also identified multiple pre-eclampsia associated genes which are
  differentially expressed in \textit{Homo sapiens} when compared to
  the other %
  \Sexpr{star.logs[,length(unique(species))-1]} species. %

  \textbf{Discussion} %
  Using the strictest filtering strategy we find a core set of only
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  genes is essential for the proper functioning of the placenta in all
  species analyzed. We also found significantly higher expression of
  the pre-eclampsia associated genes \ac{KISS1}, \ac{CRH},
  \ac{PAPPA}, and \ac{ADAM12} on the \human lineage.

  \DEW{The discussion and results seem redundant. Given the title
    shouldn't something about shape be mentioned?}

>>>>

\begin{abstract}
  \textbf{Introduction} The placenta is arguably the most anatomically
  variable organ in mammals even though its primary function is
  conserved.

@@

  \textbf{Results} We identified a set of
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  core genes which are expressed (FPKM $\ge 10$) in all eutherian placentas, %
  %
  including genes with immune-modulating properties (\ac{ANXA2},
  \ac{ANXA1}, \ac{S100A11}, \ac{S100A10}, and \ac{LGALS1}), cell-cell
  interactions (\ac{LAMC1}, \ac{LUM}, and \ac{LGALS1}),
  invasion (\ac{GRB2} and \ac{RALB}) and syncytialization
  (\ac{ANXA5} and \ac{ANXA1}). %
  We also identified multiple pre-eclampsia associated genes which are
  differentially expressed in \textit{Homo sapiens} when compared to
  the other %
  \Sexpr{star.logs[,length(unique(species))-1]} species. %
  %
  Multiple genes are significantly associated with placenta
  morphology, including \ac{EREG} and \ac{WNT5A} which are both
  associated with placental shape.

  \textbf{Discussion} %
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  genes are important for the core functions of the placenta in all
  eutherian species analyzed. The molecular functions and pathways
  enriched in the core placenta align with the evolutionarily
  conserved functionality of the placenta.

 % \DEW{The discussion and results seem redundant. Given the title
 %   shouldn't something about shape be mentioned?}
